Motional Ground State Cooling Outside the Lamb-Dicke Regime

### 2017-07-20 paper.tex
Implement more suggestions

\begin{abstract}
  We report Raman sideband cooling of a single sodium atom to its three-dimensional
  motional ground state in an optical tweezer.
  Despite having a very large Lamb-Dicke parameter, high initial temperature, and
  large differential AC Stark shifts in the excited state,
  we achieve a ground state preparation fidelity of $77(4)$\% after $100$ ms of cooling.
  Our technique includes addressing very high-order sidebands and
  fast modulation of the optical tweezer trap.
  We demonstrate that Raman sideband cooling to the 3D motional ground state is applicable
  outside the Lamb-Dicke regime, for example in
  systems where tight confinement and low initial temperature are difficult to realize.
  This is particularly relevant for systems which are challenging to laser-cool,
  such as molecules and exotic atoms,
  and opens up the possibility to gain quantum motional control of these systems.
\end{abstract}

@@

\begin{figure*}
  \includegraphics[height=5.2cm]{imgs/Na_RSC_schematic.pdf}
  \includegraphics[height=4.5cm]{sequence.pdf}
  \caption{(A) Energy levels and schematic of Raman sideband cooling.
    The Raman transitions have a one photon detuning $\Delta=25$ GHz from the D2 line.
    We use D1 light with $\sigma^-$ polarization to repump atoms out of $|F=2,m_F=-1\rangle$
    state to minimize heating on the atom in the $|F=2,m_F=-2\rangle$ state
    (B) Geometry and polarizations of the Raman and optical pumping beams relative to the
    optical tweezer and bias magnetic field.
    (C) Schematic of the cooling sequence. The tweezer switches at 3 MHz to
    reduce light shifts during optical pumping. Each cooling cycle consists of $8$ pulses.
    The four axial pulses are addressing two neighboring cooling orders.
    The two pulses in each radial direction are either addressing two neighboring cooling orders
    or having different length on the first order when most of the population are below $n=3$
    towards the end of the cooling sequence.
    All of the Raman pulses use Blackman pulse envelopes to reduce the off-resonant
    coupling to carrier and heating sidebands.
    \label{f-setup}}
\end{figure*}
\begin{figure}[b]
  \includegraphics[width=8.5cm]{imgs/fig2_raman_op.pdf}
  \caption{Optical pumping state changing probability and Raman matrix element for axial direction.
    The range plotted covers $95$\% of the initial thermal distribution.
    (A) Axial motional state distribution after one optical pumping cycle
    for different initial $n$. Due to the large Lamb-Dicke parameter,
    there is a high probability of $n$ changing.
    (B) Matrix elements for Raman transition in the axial direction showing deviation from
    $\sqrt{n}$ scaling and multiple minima for different sideband orders.
    ($k$ is the difference in wave vector of the Raman beams,
    $r$ is the coordinate in axial direction).
    \label{f-ld}}
\end{figure}
\begin{figure*}
  \includegraphics[height=4.2cm]{imgs/spectrum_r.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_r3_0.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_r3_p1.pdf}
  \caption{(A) Radial Raman sideband spectrum of first order heating, first order cooling and
    second order cooling before and after Raman sideband cooling.
    (B,C) Rabi flopping on axis 3 (B) carrier and (C) first order heating sideband
    before (red) and after (blue) cooling.
    Solid lines in (B) and (C) are theoretical calculation of the Rabi flopping.
    The blue lines correspond to a ground state probability of $93$\% after cooling and
    the red lines correspond to a thermal distribution of $70$ $\mu$K before cooling.
    \label{f-radial}}
\end{figure*}
\begin{figure*}
  \includegraphics[height=4.2cm]{imgs/spectrum_a1.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_a1_0.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_a1_p1.pdf}
  \caption{(A) Axial Raman sideband spectrum from first order heating to eighth order cooling
    before and after Raman sideband cooling.
    The data for the second and higher orders of cooling sidebands are taken with 150 $\mu$s
    pulse time and the rest are taken with 125 $\mu$s pulse time.
    (B,C) Rabi flopping on axial (B) carrier and (C) first order heating sideband
    before (red) and after (blue) cooling.
    Solid lines in (B) and (C) are theoretical calculation of the Rabi flopping.
    The blue and green lines corresponds to a ground state probability of $92$\% after cooling and
    the red lines corresponds to a thermal distribution of $70$ $\mu$K before cooling.
    The blue lines do not take into account the effect of decoherence due to resonance
    fluctuation. By comparing it to the green line in (C), which includes a $3$ kHz fluctuation,
    we can clearly see that this effect is the strongest for the post-cooling data on
    the axial heating sideband where the Rabi frequency is the lowest.
    \label{f-axial}}
\end{figure*}

@@

Fortunately, the large LD parameters also provide opportunities to overcome OP heating.
The Raman transitions in our configuration have
$\{\eta^R_{1},\eta^R_{2},\eta^R_{3}\} = \{0.40(1), 0.341(2), 0.291(1)\}$,
which allow strong coupling to higher order sidebands for atoms in higher motional states
as calculated in Figure~\ref{f-ld}B.
To offset heating from OP initially, higher-order Raman cooling sidebands can be utilized
to remove more motional quanta in a single cooling pulse.
Since the coupling strengths of different orders do not reach minima for the same pulse duration,
using multiple orders of cooling sidebands avoids accumulations of population
near the coupling minima.

@@

There are two additional important challenges in cooling single sodium atoms.
First, the high initial temperature populates high motional states,
meaning that the atoms sample the anharmonicity of the trap away from the center.
In order to address this, we drive the sidebands with a large Rabi frequency
and short pulses to Fourier broaden the spectrum coverage, and use Blackman pulse envelopes
to avoid unwanted coupling to the carrier and heating sidebands~\cite{Kasevich1992}.
The anharmonicity in the radial directions is calculated to be 1kHz/n, meaning that a
Rabi frequency of at least $15$ kHz is needed to couple to $99$\% population up to $n=15$.
Second, the tweezer trap causes a large AC Stark shift (as large as $300$ MHz)
in the excited state. This creates a large, position-dependent OP detuning
and mixes the excited-state hyperfine levels, which reduces the optical pumping fidelity.
We address this by switching the trapping light at 3 MHz during the whole cooling sequence,
similar to our loading and imaging process~\cite{Hutzler2017-LightShifts},
while leaving OP on with constant intensity.
Due to the large light shift, the OP is effectively off whenever the trap light switches back on.
Since the atom can only be addressed by the optical pumping light when the trap light is off,
the effect of light shifts on OP fidelity is suppressed.

@@

% Define \eta_R
% Lamb-Dicke parameter square explain
% why not just molecules instead of polar molecules
% entropy
% use different line shape for rabi flopping
% mention axis definitions
% heating from trap -> decrease of ground state procetage (maybe add temprature too)
% Differetial AC starck shift **between excited states and ground states**?
% at the same time, the atom is cooled via PGC instead of "the same imaging light is used to"
% 1 sentence to explain why anisotropy?
% is strobed at 3MHz instead of is switching at...
% Cite Na loading paper in caption?
% Make sure the caption for the before cooling axial rabi flopping mentions decoherence
% Mention after cooling is non-thermal?
% Make it clear that we cool for 6-15 cycles each time before going to the next set of orders in axial

>>>>

\begin{abstract}
  We report Raman sideband cooling of a single sodium atom to its three-dimensional
  motional ground state in an optical tweezer.
  Despite having a very large Lamb-Dicke parameter, high initial temperature, and
  large differential AC Stark shifts between the excited state and the ground state,
  we achieve a ground state preparation fidelity of $77(4)$\% after $100$ ms of cooling.
  Our technique includes addressing very high-order sidebands and
  fast modulation of the optical tweezer trap.
  We demonstrate that Raman sideband cooling to the 3D motional ground state is applicable
  outside the Lamb-Dicke regime, for example in
  systems where tight confinement and low initial temperature are difficult to realize.
  This is particularly relevant for systems which are challenging to laser-cool,
  such as molecules and exotic atoms,
  and opens up the possibility to gain quantum motional control of these systems.
\end{abstract}

@@

\begin{figure*}
  \includegraphics[height=5.2cm]{imgs/Na_RSC_schematic.pdf}
  \includegraphics[height=4.5cm]{sequence.pdf}
  \caption{(A) Energy levels and schematic of Raman sideband cooling.
    The Raman transitions have a one photon detuning $\Delta=25$ GHz from the D2 line.
    We use D1 light with $\sigma^-$ polarization to repump atoms out of $|F=2,m_F=-1\rangle$
    state to minimize heating on the atom in the $|F=2,m_F=-2\rangle$ state
    (B) Geometry and polarizations of the Raman and optical pumping beams relative to the
    optical tweezer and bias magnetic field.
    (C) Schematic of the cooling sequence. The tweezer is strobed at 3 MHz to
    reduce light shifts during optical pumping~\cite{Hutzler2017-LightShifts}.
    Each cooling cycle consists of $8$ pulses.
    The four axial pulses are addressing two neighboring cooling orders.
    The two pulses in each radial direction are either addressing two neighboring cooling orders
    or having different length on the first order when most of the population are below $n=3$
    towards the end of the cooling sequence.
    All of the Raman pulses use Blackman pulse envelopes to reduce the off-resonant
    coupling to carrier and heating sidebands.
    \label{f-setup}}
\end{figure*}
\begin{figure}[b]
  \includegraphics[width=8.5cm]{imgs/fig2_raman_op.pdf}
  \caption{Optical pumping state changing probability and Raman matrix element for axial direction.
    The range plotted covers $95$\% of the initial thermal distribution.
    (A) Axial motional state distribution after one optical pumping cycle
    for different initial $n$. Due to the large Lamb-Dicke parameter,
    there is a high probability of $n$ changing.
    (B) Matrix elements for Raman transition in the axial direction showing deviation from
    $\sqrt{n}$ scaling and multiple minima for different sideband orders.
    ($k$ is the difference in wave vector of the Raman beams,
    $r$ is the coordinate in axial direction).
    \label{f-ld}}
\end{figure}
\begin{figure*}
  \includegraphics[height=4.2cm]{imgs/spectrum_r.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_r3_0.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_r3_p1.pdf}
  \caption{(A) Radial Raman sideband spectrum of first order heating, first order cooling and
    second order cooling before and after Raman sideband cooling.
    (B,C) Rabi flopping on axis 3 (B) carrier and (C) first order heating sideband
    before (red) and after (blue) cooling.
    Solid lines in (B) and (C) are theoretical calculation of the Rabi flopping.
    The blue lines correspond to a ground state probability of $93$\% after cooling and
    the red lines correspond to a thermal distribution of $70$ $\mu$K before cooling.
    \label{f-radial}}
\end{figure*}
\begin{figure*}
  \includegraphics[height=4.2cm]{imgs/spectrum_a1.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_a1_0.pdf}
  \includegraphics[height=4.2cm]{imgs/rabi_flop_a1_p1.pdf}
  \caption{(A) Axial Raman sideband spectrum from first order heating to eighth order cooling
    before and after Raman sideband cooling.
    The data for the second and higher orders of cooling sidebands are taken with 150 $\mu$s
    pulse time and the rest are taken with 125 $\mu$s pulse time.
    (B,C) Rabi flopping on axial (B) carrier and (C) first order heating sideband
    before (red) and after (blue) cooling.
    Solid lines in (B) and (C) are theoretical calculation of the Rabi flopping.
    The blue and green lines corresponds to a ground state probability of $92$\% after cooling and
    the red lines corresponds to a thermal distribution of $70$ $\mu$K before cooling.
    The blue lines do not take into account the effect of decoherence due to resonance
    fluctuation. By comparing it to the green line in (C), which includes a $3$ kHz fluctuation,
    we can clearly see that this effect is the strongest for the post-cooling data on
    the axial heating sideband where the Rabi frequency is the lowest.
    \label{f-axial}}
\end{figure*}

@@

Fortunately, the large LD parameters also provide opportunities to overcome OP heating.
The Raman transitions in our configuration have LD parameters of
$\{\eta^R_{1},\eta^R_{2},\eta^R_{3}\} = \{0.40(1), 0.341(2), 0.291(1)\}$,
which allow strong coupling to higher order sidebands for atoms in higher motional states
as calculated in Figure~\ref{f-ld}B.
The LD parameter for Raman transitions is defined as $\eta^R\equiv\Delta k z_0$ similar to OP
where $\Delta k$ being the wave vector difference
between the two beams that drives the Raman transition.
To offset heating from OP initially, higher-order Raman cooling sidebands can be utilized
to remove more motional quanta in a single cooling pulse.
Since the coupling strengths of different orders do not reach minima for the same pulse duration,
using multiple orders of cooling sidebands avoids accumulations of population
near the coupling minima.

@@

There are two additional important challenges in cooling single sodium atoms.
First, the high initial temperature populates high motional states,
meaning that the atoms sample the anharmonicity of the trap away from the center.
In order to address this, we drive the sidebands with a large Rabi frequency
and short pulses to Fourier broaden the spectrum coverage, and use Blackman pulse envelopes
to avoid unwanted coupling to the carrier and heating sidebands~\cite{Kasevich1992}.
The anharmonicity in the radial directions is calculated to be 1kHz/n, meaning that a
Rabi frequency of at least $15$ kHz is needed to couple to $99$\% population up to $n=15$.
Second, the tweezer trap causes a large AC Stark shift (as large as $300$ MHz)
in the excited state. This creates a large, position-dependent OP detuning
and mixes the excited-state hyperfine levels, which reduces the optical pumping fidelity.
We address this by strobing the trapping light at 3 MHz during the whole cooling sequence,
similar to our loading and imaging process~\cite{Hutzler2017-LightShifts},
while leaving OP on with constant intensity.
Due to the large light shift, the OP is effectively off whenever the trap light switches back on.
Since the atom can only be addressed by the optical pumping light when the trap light is off,
the effect of light shifts on OP fidelity is suppressed.

@@

% Lamb-Dicke parameter square explain
% why not just molecules instead of polar molecules
% entropy
% use different line shape for rabi flopping
% mention axis definitions
% heating from trap -> decrease of ground state procetage (maybe add temprature too)
% at the same time, the atom is cooled via PGC instead of "the same imaging light is used to"
% 1 sentence to explain why anisotropy?
% Make sure the caption for the before cooling axial rabi flopping mentions decoherence
% Mention after cooling is non-thermal?
% Make it clear that we cool for 6-15 cycles each time before going to the next set of orders in axial
